---
title: "The Influence of Net Migration, Unemployment, and Inflation on Economic Growth"
author: "LUIS HEREDIA"
date: "2024-12-12"
output:
  html_document: default
  pdf_document: default
---

## Summary

Load necessary libraries, otherwise the code won't run. If at any point you find a bug, please let us know. The code is presented in chunks, therefore there will be different output based on previous functions. Run one after another.

## Shortcut
To see the data combined use the command View(combined_data_wide) on the command line interface. The first 4 columns will be composed of the observed and explanatory variables.Besides, the ggsave, command after each plot is executed, since it was used locally. 




```{r}
library(ggplot2)
library(dplyr)
library(readr)
library(purrr)
library(tidyr)
library(broom)
library(modelr)
library(readxl)
```

```{r}
# Load datasets. These are the data used for the model and graphs 
inflation <- read.csv("inflation.csv")
unrate <- read.csv("unrate.csv")
real_gdp <- read.csv("GDPC1.csv") |>  mutate(DATE = as.Date(DATE))
world_net_migration<- read_excel("Net_Immigration_world_data.xlsx", sheet = "Sheet1", skip = 3)
```

```{r}
#Tidying world_net_immigration data, from the World Bank, to get only values for the US
US_Net_migration <- world_net_migration |>  filter(`Country Name` == "United States")|> select(-c(`Country Code`, `Indicator Name`, `Indicator Code`)) |> #Remove metadata, what we will not use
pivot_longer(cols = starts_with("19") | starts_with("20"), 
               names_to = "Year", values_to = "NetMigration") |> arrange(as.numeric(Year)) |> mutate("DATE"=Year, PercentChange = (NetMigration - lag(NetMigration)) / lag(NetMigration) * 100) |> mutate(DATE = as.Date(paste0(DATE, "-01-01")))|> slice(-1) |> select(DATE,PercentChange)
US_Net_migration

```
```{r}

# US NET MIGRATION DATA VISUALIZATION
US_Net_migration_plot <- ggplot(US_Net_migration, aes(x = DATE, y = PercentChange)) +
  geom_point(color = "blue", alpha = 0.7, size = 2) +  
  geom_line(color = "darkgreen", size = 1) +          
  scale_x_date(
    date_labels = "%Y",
    breaks = c(seq(as.Date("1961-01-01"), as.Date("2020-01-01"), by = "5 years"),
               as.Date("2023-01-01"))
  ) +
  geom_smooth(method = "loess", color = "red", linetype = "dashed", size = 1) +
  labs(title = "Net Migration Growth: 1961 to 2023",
       x = "Year", y = "Net Migration (% Change)") +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(hjust = 0.5),  # Center title
    axis.title.x = element_text(margin = margin(t = 10)),  # x space title
    axis.title.y = element_text(margin = margin(r = 10)),  # y space title 
    axis.text.x = element_text(angle = 45, hjust = 1)  # Rotation
  )

US_Net_migration_plot
#You may want to save the code locally, change the file location if needed, then run the code which is given below.
#ggsave(
#  filename = "C:/Users/luisc/Desktop/CODE_STATISTICS_ECON57/IMAGES/netmigration_plot_visualization.jpg", 
#  plot = US_Net_migration_plot,
#  width = 12, 
 # height = 8, 
 # dpi = 600
#)
```


```{r}
#Date format
inflation <- inflation |>
  mutate(DATE = as.Date(DATE))

# inflation_plot, DATA VISUALIZATION
inflation_plot <- ggplot(inflation, aes(x = DATE, y = FPCPITOTLZGUSA)) +
  geom_point(color = "blue", alpha = 0.7, size = 2) +  
  geom_line(color = "darkgreen", size = 1) +          
  scale_x_date(
    date_labels = "%Y",
    breaks = c(seq(as.Date("1961-01-01"), as.Date("2020-01-01"), by = "5 years"),
               as.Date("2023-01-01")) 
  ) +
  geom_smooth(method = "loess", color = "red", linetype = "dashed", size = 1) +
  labs(title = "Inflation: 1961 to 2023",
       x = "Year", y = "Inflation (%)") +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(hjust = 0.5),  # Center title
    axis.title.x = element_text(margin = margin(t = 10)),  # x space title
    axis.title.y = element_text(margin = margin(r = 10)),  # y space title 
    axis.text.x = element_text(angle = 45, hjust = 1)  # Rotation
  )

inflation_plot

#ggsave(
#  filename = "C:/Users/luisc/Desktop/CODE_STATISTICS_ECON57/IMAGES/inflation_plot_visualization.jpg", 
#  plot = inflation_plot,
#  width = 12, 
  #height = 8, 
 # dpi = 600
#)

```

```{r}

#Date format
unrate <- unrate |>
  mutate(DATE = as.Date(DATE))

# DATA VISUALIZATION FOR UNEMPLOYMENT RATE

unrate_plot <- ggplot(unrate, aes(x = DATE, y = UNRATE)) +
  geom_point(color = "blue", alpha = 0.7, size = 2) +  
  geom_line(color = "darkgreen", size = 1) +          
  scale_x_date(
    date_labels = "%Y",
    breaks = c(seq(as.Date("1961-01-01"), as.Date("2020-01-01"), by = "5 years"),
               as.Date("2023-01-01")) 
  ) +
  geom_smooth(method = "loess", color = "red", linetype = "dashed", size = 1) +
  labs(title = "Unemployment Rate Growth: 1961 to 2023",
       x = "Year", y = "Unemployment Rate Growth (%)") +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(hjust = 0.5),  
    axis.title.x = element_text(margin = margin(t = 10)),  
    axis.title.y = element_text(margin = margin(r = 10)),  
    axis.text.x = element_text(angle = 45, hjust = 1) 
  )
unrate_plot

#ggsave(
 # filename = "C:/Users/luisc/Desktop/CODE_STATISTICS_ECON57/IMAGES/unrate_plot_visualization.jpg", 
  #plot = unrate_plot,
 # width = 12, 
  #height = 8, 
  #dpi = 600
#)

```

```{r}


#Date format
real_gdp<- real_gdp |>
  mutate(DATE = as.Date(DATE))

# Real GDP, data visualization

real_gdp_plot <- ggplot(real_gdp, aes(x = DATE, y= GDPC1_PC1)) +
  geom_point(color = "blue", alpha = 0.7, size = 2) +  
  geom_line(color = "darkgreen", size = 1) +          
  scale_x_date(
    date_labels = "%Y",
    breaks = c(seq(as.Date("1961-01-01"), as.Date("2020-01-01"), by = "5 years"),
               as.Date("2023-01-01")) 
  ) +
  geom_smooth(method = "loess", color = "red", linetype = "dashed", size = 1) +
  labs(title = "Real GDP Growth: 1961 to 2023",
       x = "Year", y = "GDP Growth (%)") +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(hjust = 0.5),  # Center title
    axis.title.x = element_text(margin = margin(t = 10)), 
    axis.title.y = element_text(margin = margin(r = 10)),  
    axis.text.x = element_text(angle = 45, hjust = 1)  
  )
real_gdp_plot

#ggsave(
#  filename = "C:/Users/luisc/Desktop/CODE_STATISTICS_ECON57/IMAGES/real_gdp_plot_visualization.jpg", 
#  plot = real_gdp_plot,
#  width = 12, 
#  height = 8, 
# dpi = 600
#)

```

```{r}
# Combine all data, including US Net Migration Percent Change
combined_data <- real_gdp |>
  select(DATE, GDP_Growth = GDPC1_PC1) |>
  left_join(unrate |> select(DATE, Unemployment_Rate = UNRATE), by = "DATE") |>
  left_join(inflation |> select(DATE, inflation = FPCPITOTLZGUSA), by = "DATE") |>
  left_join(US_Net_migration |> rename(Percentage_Net_Migration_Change = PercentChange), by = "DATE") |>
  pivot_longer(
    cols = c(GDP_Growth, Unemployment_Rate, inflation, Percentage_Net_Migration_Change),
    names_to = "Indicator",
    values_to = "Value"
  )

# Checking ranges of the indicators, including Net Migration Change
combined_data_range <- combined_data |>
  group_by(Indicator) |>
  summarize(
    Min = min(Value, na.rm = TRUE),
    Max = max(Value, na.rm = TRUE)
  )

combined_data_range

```



```{r}

combined_data <- combined_data |>
  group_by(Indicator) |>
  mutate(Normalized_Value = (Value - mean(Value, na.rm = TRUE)) / sd(Value, na.rm = TRUE)) |>
  ungroup()
# Add recession periods, including COVID-19 recession
recession_periods <- data.frame(
  start = as.Date(c("1973-01-01", "1980-01-01", "1990-07-01", "2008-01-01", "2020-02-01")),
  end = as.Date(c("1975-01-01", "1982-11-01", "1991-03-01", "2009-06-01", "2020-04-01")),
  label = c("1973-75", "1980-82", "1990-91", "2008-09", "COVID-19 (2020)")
)

# Highlight high and low points (example for GDP Growth)
highlight_points <- combined_data |>
  filter(Indicator == "GDP Growth" & (Normalized_Value == max(Normalized_Value) | 
                                      Normalized_Value == min(Normalized_Value)))

# Enhanced plot with annotations
overlay_plot_normalized <- ggplot(combined_data, aes(x = DATE, y = Normalized_Value, color = Indicator)) +
  # Add lines
  geom_line(size = 0.8, alpha = 0.8) +
  # Add shaded recession periods
  geom_rect(data = recession_periods,
            aes(xmin = start, xmax = end, ymin = -Inf, ymax = Inf),
            fill = "gray", alpha = 0.2, inherit.aes = FALSE) +
  # Annotate recession periods
  geom_text(data = recession_periods,
            aes(x = start + (end - start) / 2, y = 4.5, label = label),  # Position labels above the lines
            color = "black", size = 4, inherit.aes = FALSE) +
  # Highlight high and low points
  geom_point(data = highlight_points, aes(x = DATE, y = Normalized_Value), 
             color = "black", size = 3) +
  # Annotate high and low points
  geom_text(data = highlight_points, aes(x = DATE, y = Normalized_Value, 
                                         label = round(Normalized_Value, 2)), 
            color = "black", hjust = -0.2, vjust = -0.5, size = 4) +
  # Add labels and formatting
  labs(
    title = "Economic and Migration Indicators: 1961 to 2023",
    x = "Year",
    y = "Z-Score (Normalized Values)",
    color = "Indicators"
  ) +
  scale_color_manual(
    values = c("blue", "red", "green", "purple"),
    labels = c("Inflation", "GDP Growth", "Unemployment Rate", "Net Migration Change")
  ) +
  scale_x_date(
    date_labels = "%Y",
    breaks = c(seq(as.Date("1961-01-01"), as.Date("2020-01-01"), by = "5 years"),
               as.Date("2023-01-01"))
  ) +
  coord_cartesian(ylim = c(-3, 5)) +
  theme_minimal(base_size = 16) +
  theme(
    plot.title = element_text(hjust = 0.5, size = 18),
    axis.text.x = element_text(angle = 45, hjust = 1, size = 12),
    axis.text.y = element_text(size = 12),
    legend.position = "top",
    legend.text = element_text(size = 12),
    legend.title = element_text(size = 14),
    panel.grid.major = element_line(size = 0.2, color = "gray80"),
    panel.grid.minor = element_blank()
  )

# Display the enhanced plot
overlay_plot_normalized

#ggsave(
#  filename = "C:/Users/luisc/Desktop/CODE_STATISTICS_ECON57/IMAGES/normalized_data_all_indicators.jpg", 
# plot = overlay_plot_normalized,
 # width = 12, 
#  height = 8, 
 # dpi = 600
#)

```


```{r}


combined_data_wide <- real_gdp |>
  select(DATE, GDP_Growth = GDPC1_PC1) |>
  left_join(unrate |> select(DATE, Unemployment_Rate = UNRATE), by = "DATE") |>
  left_join(inflation |> select(DATE, inflation = FPCPITOTLZGUSA), by = "DATE") |>
  left_join(US_Net_migration |> rename(Net_Migration = PercentChange), by = "DATE") # Add Net Migration

# Net Migration is the third column
combined_data_wide <- combined_data_wide |>
  select(DATE, GDP_Growth, Net_Migration, Unemployment_Rate, inflation)

# Check structure
head(combined_data_wide)

```

```{r}
#Summary statistics

library(psych)
describe(combined_data_wide) |> slice(-1) |> select(-1) |> mutate("observations"=n)|> select(-trimmed,-mad,-range,-n) |>  select(-skew,-kurtosis,-se)|> slice(1:4)

```

```{r}
# Simple regression
simple_model <- lm(GDP_Growth ~ Unemployment_Rate, data = combined_data_wide)

# Summary of the simple regression model
summary(simple_model)

```

```{r}
data_predictors <- combined_data_wide |> data_grid(Unemployment_Rate)|> add_predictions(simple_model)
data_predictors
```

```{r}
residuals_unemployment <-combined_data_wide |> select(GDP_Growth,Unemployment_Rate)|> add_residuals(simple_model)
residuals_unemployment
```



```{r}
#This data if for unemployment residuals, you can skeep this part
joined_predictors_residuals<- data_predictors|> left_join(residuals_unemployment)|> select(Unemployment_Rate,GDP_Growth, pred, resid)
joined_predictors_residuals
                                                      
```



```{r}
residuals_unemployment_density <-
residuals_unemployment |> 
  ggplot(aes(x = resid)) +
  
  geom_density(fill = "blue", alpha = 0.5) +  # Density plot
  geom_vline(xintercept = 0, linetype = "dashed", color = "red", size = 1) +  # Line at 0
  labs(
    title = "Density plot of residuals for unemployment",
    x = "Residuals",
    y = "Density"
  ) + 
  theme_minimal(base_size = 14) 

residuals_unemployment_density


#ggsave(
# filename = "C:/Users/luisc/Desktop/CODE_STATISTICS_ECON57/IMAGES/residuals_uemployment_density.jpg", 
  #plot = residuals_unemployment_density,
  #width = 12, 
 # height = 8, 
  #dpi = 600
#)

  

```

```{r}
# Extract coefficients from the simple regression model
coefficients <- coef(simple_model)
equation <- paste0(
  "y = ", round(coefficients[1], 2), 
  " + ", round(coefficients[2], 2), "x"
)

# observed, predicted data, and equation
plot_predictors <- ggplot() +
  # Observed points
  geom_point(data = combined_data_wide, aes(Unemployment_Rate, GDP_Growth), 
             color = "blue", alpha = 0.7, size = 3) +
  # Predicted points
  geom_point(data = data_predictors, aes(Unemployment_Rate, pred), 
             color = "red", alpha = 0.7, size = 3) +
  # Regression line
  geom_smooth(data = combined_data_wide, aes(Unemployment_Rate, GDP_Growth), 
              method = "lm", color = "darkblue", se = FALSE) +
  # Add regression equation
  annotate("text", 
           x = min(combined_data_wide$Unemployment_Rate) + 1, 
           y = max(combined_data_wide$GDP_Growth) - 1, 
           label = equation, 
           hjust = 0, size = 5, color = "darkred") +
  # Labels and title
  labs(
    title = "GDP Growth vs Unemployment Rate",
    x = "Unemployment Rate (%)",
    y = "GDP Growth (%)",
    caption = "Blue: Observed Data | Red: Predicted Data"
  ) +
  # Minimal theme
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    axis.title = element_text(face = "bold"),
    legend.position = "none"
  )
plot_predictors

#ggsave(
# filename = "C:/Users/luisc/Desktop/CODE_STATISTICS_ECON57/IMAGES/UR_REGRESSION.jpg", 
#plot = plot_predictors,
# width = 12, 
# height = 8, 
# dpi = 600
#)


```

```{r}

# Simple regression
simple_model_net_migration <- lm(GDP_Growth ~ Net_Migration, data = combined_data_wide)

# Summary of the simple regression model
summary(simple_model_net_migration)

```

```{r}
# Extract coefficients from the model
coefficients <- coef(simple_model_net_migration)
equation <- paste0(
  "y = ", round(coefficients[1], 2), 
  " + ", round(coefficients[2], 2), "x"
)

# Add predicted values for Net Migration
combined_data_wide <- combined_data_wide |> 
  mutate(predicted_net_migration = predict(simple_model_net_migration))

# Plot observed vs. predicted with the regression line
plot_net_migration <- ggplot() +
  # Observed points
  geom_point(data = combined_data_wide, aes(x = Net_Migration, y = GDP_Growth), 
             color = "blue", alpha = 0.7, size = 3) +
  # Predicted points
  geom_point(data = combined_data_wide, aes(x = Net_Migration, y = predicted_net_migration), 
             color = "red", alpha = 0.7, size = 3) +
  # Regression line
  geom_smooth(data = combined_data_wide, aes(x = Net_Migration, y = GDP_Growth), 
              method = "lm", color = "darkblue", se = FALSE, size = 1) +
  # Add regression equation
  annotate("text", 
           x = min(combined_data_wide$Net_Migration) + 0.5, 
           y = max(combined_data_wide$GDP_Growth) - 0.5, 
           label = equation, 
           hjust = 0, size = 5, color = "darkred") +
  # Labels and title
  labs(
    title = "GDP Growth vs Net Migration",
    x = "Net Migration (%)",
    y = "GDP Growth (%)",
    caption = "Blue: Observed Data | Red: Predicted Data"
  ) +
  # Minimal theme
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    axis.title = element_text(face = "bold"),
    legend.position = "none"
  )

# Display the plot
print(plot_net_migration)
#ggsave(
 # filename = "C:/Users/luisc/Desktop/CODE_STATISTICS_ECON57/IMAGES/NET_MIGRATION_REGRESSION.jpg", 
 # plot = plot_net_migration,
 # width = 12, 
 # height = 8, 
  #dpi = 600
#)

```
```{r}
# Calculate residuals for the Net Migration model
combined_data_wide <- combined_data_wide |>
  mutate(residuals_net_migration = GDP_Growth - predicted_net_migration)

# Plot the residuals using a density plot
net_migration_density<-
ggplot(combined_data_wide, aes(x = residuals_net_migration)) +
  geom_density(fill = "blue", alpha = 0.5) +  # Density plot
  geom_vline(xintercept = 0, linetype = "dashed", color = "red", size = 1) +  # Line at 0
  labs(
    title = "Density Plot of Residuals for Net Migration Model",
    x = "Residuals (Net Migration Model)",
    y = "Density"
  ) +
  theme_minimal(base_size = 14)
net_migration_density
#ggsave(
 # filename = "C:/Users/luisc/Desktop/CODE_STATISTICS_ECON57/IMAGES/NET_MIGRATION_DENSITY.jpg", 
 # plot = net_migration_density,
 # width = 12, 
  #height = 8, 
  #dpi = 600
#)

```

```{r}
# Simple regression for inflation
simple_model_inflation <- lm(GDP_Growth ~ inflation, data = combined_data_wide)

# Summary of the simple regression model
summary(simple_model_inflation)


```

```{r}
# Extract coefficients from the model
coefficients_cpi <- coef(simple_model_inflation)
equation_cpi <- paste0(
  "y = ", round(coefficients_cpi[1], 2), 
  " + ", round(coefficients_cpi[2], 2), "x"
)

# Add predicted values for CPI Growth
combined_data_wide <- combined_data_wide |>
  mutate(predicted_inflation = predict(simple_model_inflation))

# Plot observed vs. predicted with the regression line
plot_inflation <- ggplot() +
  # Observed points
  geom_point(data = combined_data_wide, aes(x = inflation, y = GDP_Growth), 
             color = "blue", alpha = 0.7, size = 3) +
  # Predicted points
  geom_point(data = combined_data_wide, aes(x = inflation, y = predicted_inflation), 
             color = "red", alpha = 0.7, size = 3) +
  # Regression line
  geom_smooth(data = combined_data_wide, aes(x = inflation, y = GDP_Growth), 
              method = "lm", color = "darkblue", se = FALSE, size = 1) +
  # Add regression equation
  annotate(
    "text", 
    x = min(combined_data_wide$inflation) + 0.5, 
    y = max(combined_data_wide$GDP_Growth) - 0.5, 
    label = equation_cpi, 
    hjust = 0, size = 5, color = "darkred"
  ) +
  # Labels and title
  labs(
    title = "GDP Growth vs Inflation",
    x = "Inflation (%)",
    y = "GDP Growth (%)",
    caption = "Blue: Observed Data | Red: Predicted Data"
  ) +
  # Minimal theme
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    axis.title = element_text(face = "bold"),
    legend.position = "none"
  )

# Display the plot
print(plot_inflation)
#ggsave(
#  filename = "C:/Users/luisc/Desktop/CODE_STATISTICS_ECON57/IMAGES/plot_regression_inflation.jpg", 
#  plot = plot_inflation,
#  width = 12, 
#  height = 8, 
#  dpi = 600
#)


```


```{r}

# Calculate residuals for the CPI Growth model
combined_data_wide <- combined_data_wide |>
  mutate(residuals_inflation = GDP_Growth - predicted_inflation)

# Plot the residuals using a density plot
inflation_density<-
ggplot(combined_data_wide, aes(x = residuals_inflation)) +
  geom_density(fill = "blue", alpha = 0.5) +  # Density plot
  geom_vline(xintercept = 0, linetype = "dashed", color = "red", size = 1) +  # Line at 0
  labs(
    title = "Density Plot of Residuals for Inflation",
    x = "Residuals (inflation Model)",
    y = "Density"
  ) +
  theme_minimal(base_size = 14)
inflation_density

#ggsave(
 # filename = "C:/Users/luisc/Desktop/CODE_STATISTICS_ECON57/IMAGES/inflation_density.jpg", 
 # plot = inflation_density,
 # width = 12, 
 # height = 8, 
 # dpi = 600
#)


```

```{r}

# Multiple regression model with CPI Growth, Unemployment Rate, and Net Migration
multi_model <- lm(GDP_Growth ~ Net_Migration + Unemployment_Rate + inflation, data = combined_data_wide)

# Add predicted values from the multiple model to the dataset
combined_data_wide <- combined_data_wide |>
  mutate(Predicted_GDP = predict(multi_model))  # Column for multiple model predictions

head(combined_data_wide)

```




```{r}
# Summary of the multi regression model
summary(multi_model)

```


```{r}
# Ensure no missing data
combined_data_wide <- combined_data_wide |>
  filter(!is.na(Net_Migration), !is.na(Unemployment_Rate), !is.na(inflation))

# Add predicted values and residuals from the multivariable model
combined_data_wide <- combined_data_wide |>
  mutate(
    Predicted_GDP_Multimodel = predict(multi_model),  # Predicted GDP from the model
    residuals = GDP_Growth - Predicted_GDP_Multimodel  # Residuals (Observed - Predicted)
  )

# Identify outliers by residuals (top 5% largest residuals)
outliers <- combined_data_wide |>
  filter(abs(residuals) > quantile(abs(residuals), 0.95)) |>
  mutate(outlier_id = row_number())  # Assign unique IDs to outliers

# Extract coefficients and create the equation
coefficients <- coef(multi_model)
equation <- paste0(
  "y = ", round(coefficients[1], 2), 
  " + ", round(coefficients[2], 4), " * Net_Migration",
  " + ", round(coefficients[3], 4), " * Unemployment_Rate",
  " + ", round(coefficients[4], 4), " * Inflation"
)

# Plot observed vs. predicted with explicit differentiation
multi_plot <- ggplot(combined_data_wide) +
  # Observed values (y-axis)
  geom_point(aes(x = Predicted_GDP_Multimodel, y = GDP_Growth, color = "Observed"), 
             alpha = 0.7, size = 3) +
  # Predicted values (x-axis)
  geom_point(aes(x = Predicted_GDP_Multimodel, y = Predicted_GDP_Multimodel, color = "Predicted"), 
             alpha = 0.7, size = 3) +
  # Residual lines with blue color
  geom_segment(aes(x = Predicted_GDP_Multimodel, y = Predicted_GDP_Multimodel, 
                   xend = Predicted_GDP_Multimodel, yend = GDP_Growth), 
               color = "darkblue", linetype = "dotted", alpha = 0.7) +
  # Perfect fit line (45-degree)
  geom_abline(slope = 1, intercept = 0, color = "black", linetype = "dashed", size = 1) +
  # Regression equation
  annotate("text", 
           x = min(combined_data_wide$Predicted_GDP_Multimodel) + 0.5, 
           y = max(combined_data_wide$GDP_Growth) - 1, 
           label = equation, 
           hjust = 0, size = 3, color = "darkred") +
  # Highlight outliers
  geom_text(data = outliers, aes(x = Predicted_GDP_Multimodel, y = GDP_Growth, label = outlier_id), 
            hjust = -0.3, vjust = -0.5, color = "darkred", size = 3) +
  # Color scale for observed and predicted points
  scale_color_manual(values = c("Observed" = "orange", "Predicted" = "purple"), name = "Point Type") +
  # Labels and title
  labs(
    title = "Multivariable Regression: Observed vs. Predicted GDP Growth",
    x = "Predicted GDP Growth (Multimodel) (%)",
    y = "Observed GDP Growth (%)",
    caption = "Orange: Observed GDP Growth | Purple: Predicted GDP Growth | Blue lines: Residuals"
  ) +
  theme_minimal(base_size = 14)
# Display the plot
print(multi_plot)

#ggsave(
#  filename = "C:/Users/luisc/Desktop/CODE_STATISTICS_ECON57/IMAGES/multi_regression_plot_high_quality.jpg", 
#  plot = multi_plot,
#  width = 12, 
#  height = 8, 
# dpi = 600
#)


```


```{r}
# Add predicted GDP Growth and calculate residuals
combined_data_wide <- combined_data_wide |>
  mutate(
    Predicted_GDP = predict(multi_model),  # Predicted values
    residuals_multi_model = GDP_Growth - Predicted_GDP  # Residuals
  )

# Plot the residuals using a density plot
residual_densisty <-
ggplot(combined_data_wide, aes(x = residuals_multi_model)) +
  geom_density(fill = "blue", alpha = 0.5) +  # Density plot
  geom_vline(xintercept = 0, linetype = "dashed", color = "red", size = 1) +  # Line at 0
  labs(
    title = "Density Plot of Residuals for Multivariable Model",
    x = "Residuals (Multivariable Model)",
    y = "Density"
  ) +
  theme_minimal(base_size = 14)
residual_densisty

#ggsave(
#  filename = "C:/Users/luisc/Desktop/CODE_STATISTICS_ECON57/IMAGES/residual_multimodel.jpg", 
 # plot = residual_densisty, 
#  width = 12, 
 # height = 8, 
 # dpi = 600
#)
```
















